---
title: Problem Set 9
author: Isaac Baron
format: html
editor: visual
---

```{r}
rm(list=ls())
library(tidyverse)
library(factoextra)
library(ggthemes)
```

Part 1: Segmenting Colleges

```{r}
scale0 = function(x) {
  (x - mean(x)) / sd(x) 
}

colleges = read_csv("college.csv") %>%
  filter(state =="IN") %>%
  column_to_rownames(var = "name") %>%
  select(faculty_salary_avg, tuition) %>%
  mutate_all(.funs = scale0) 
  
```

Check for missing values

```{r}
sapply(colleges, function(x) mean(is.na(x)))
```

```{r}
IN_clusters = kmeans(colleges,
                     center = 3,
                     nstart =25)

factoextra::fviz_cluster(IN_clusters, colleges, repel = TRUE)
```

```{r}
factoextra::fviz_nbclust(colleges, kmeans, method = "wss")
```

```{r}
factoextra::fviz_nbclust(colleges, kmeans, method = "silhouette")
```

```{r}
factoextra::fviz_nbclust(colleges, kmeans, method = "gap_stat")
```

The different approaches suggest a different number of clusters.

**Elbow method:** We see elbows at 3,5,7.

**Silhouette:** Of the 10 clusters 4 has the highest value. 3 is very close as well.

**Gap Statistic:** The optimal numbers of clusters appears to be 9 and 6.

We are going to look at 4 and 6

```{r}
college_clust_4 = kmeans(colleges,
                         centers =4,
                         nstart =25)

factoextra::fviz_cluster(college_clust_4, colleges, repel = TRUE ) 
```

```{r}
college_clust_6 =kmeans(colleges,
                        centers = 6,
                        nstart = 25)

factoextra::fviz_cluster(college_clust_6, colleges, repel = TRUE)
```

Part 2: Segmenting the cereal market

```{r}
cereal = read_csv("Cereals.csv") %>%
  column_to_rownames(var = "name") %>%
  select(-mfr,-type,-weight,-shelf,-cups,-rating) %>%
  drop_na() %>%
  mutate_all(.funs = scale0)


  
```

```{r}
sapply(cereal, function(x) sum(is.na(x)))
```

```{r}
fviz_nbclust(cereal, kmeans, method = "wss")
```

6

```{r}
fviz_nbclust(cereal, kmeans, method = "silhouette")
```

6,8,10

```{r}
fviz_nbclust(cereal, kmeans, method = "gap_stat")
```

10

```{r}
cereal_clust = kmeans(cereal,
                      centers = 10,
                      nstart = 25)

cereal = cereal %>%
  mutate(cluster = cereal_clust$cluster)

centers = cereal_clust$centers %>%
  as.data.frame() %>%
  rownames_to_column(var ="cluster")

centers
```

```{r}
ggplot(centers) +
  geom_point(aes(x = scale0(sugars),
                 y = scale0(calories),
                 color = factor(cluster)),
             size = 4.5) +
  geom_vline(xintercept = 0, linetype='dashed', alpha=0.7) +
  geom_hline(yintercept = 0, linetype='dashed', alpha=0.7) + 
  labs(title = "Calories vs. Sugar by Cluster Center",
       x = "Sugars", y = "Calories" ) +
  theme_classic()
                        
```

Part 3 Segmenting the Bath soap Customers

```{r}
soap_full = read_csv("BathSoapHousehold.csv") %>%
  mutate_all(.funs = scale0) %>%
  rename(AffluenceIndex = `Affluence Index`)

soap = soap_full %>%
  select(CHILD, AffluenceIndex)
```

```{r}
fviz_nbclust(soap, kmeans, method = "wss")
```

5

```{r}
fviz_nbclust(soap, kmeans, method = "silhouette")
```

2

```{r}
fviz_nbclust(soap, kmeans, method = "gap_stat")
```

4

Elbow Method: 5

Silhouette Method: 2

Gap_stat Method: 4

We are going to look at 3 clusters

```{r}
soap_clust_3 = kmeans(soap,
                         centers =3,
                         nstart =25)

factoextra::fviz_cluster(soap_clust_3, soap, geom = c("point") ) 
```

```{r}
soap_full %>%
  mutate(cluster = soap_clust_3$cluster) %>%
  group_by(cluster) %>%
  summarise_at(vars(Value, 'Total Volume'),funs(mean))

  
```
